---
title: "Homework 2 (tm2925)"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(readxl)
```

##Problem 1

```{r read and tidy data1}
transit_data = read_csv(file = "./data1/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() 
  skimr::skim(transit_data)
```

```{r}
transit_data_select = 
  select(transit_data, line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  gather(route1:route11, key = "route", value = "route_lines") 
#
```

The transit_data set has been tidyed by selecting the variables of interest, gathering the columns for the 11 route lines, and collapsing them into a single column, and creating a new variable that converts the entry variable from a character vector into a logical vector. The transit_data_select dataset contains `r nrow(transit_data_select)` observations and `r ncol(transit_data_select)` variables.

```{r problem 1 questions}
transit_data_select %>% 
distinct(station_name, line) 
#465 distinct stations
sum(transit_data$ada, na.rm = TRUE)
#468 stations ADA compliant
```

```{r}
#proportion of station entrances / exits without vending allow entrance
transit_data_select  %>%  
  filter(vending == "NO") 

##transit_data_select$entry_logical <- as.logical(transit_data_select$entry_logical)
##mutate(entry_logical = as.logical(entry))#
```

```{r data1 reformat}
transit_data_select %>% 
filter(route_lines == "A") %>% 
#273 stations serve A train
distinct(station_name, ada = TRUE)
#56 stations that serve the A train are ADA compliant
```


# Problem 2

```{r eval=FALSE}
trash_data = readxl::read_excel(path = "./data2/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                                sheet = 1, 
                                range = cell_cols("A:N"),
                                skip = row_number(1)) %>% 
janitor::clean_names()
skimr::skim(trash_data) 

gather(trash_data, "trash_type", value = "trash_quantity", plastic_bottles:sports_balls) 
filter(trash_data, !is.na(dumpster)) %>% 
mutate(balls_integer = as.integer(sports_balls)) %>% View
```

```{r read in data2_2017}
precip_2017 = readxl::read_excel(path = "./data2/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                                sheet = 3,
                                skip = row_number(1)) %>% 
janitor::clean_names()
filter(precip_2017, !is.na(total)) 
filter(precip_2017, !is.na(month)) %>% 
mutate(year = 2017)

precip_2016 = readxl::read_excel(path = "./data2/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
                                sheet = 4,
                                skip = row_number(1)) %>% 
janitor::clean_names()
filter(precip_2016, !is.na(total)) 
filter(precip_2016, !is.na(month)) %>% 
mutate(year = 2016)

precip_data = left_join(precip_2017, precip_2016, by = "month")
colnames(precip_data)[2] = "2017"
colnames(precip_data)[3] = "2016"
```

The Trash dataset describes the amount, volume and type of trash collected by date. It
contains `r nrow(trash_data)` observations and `r ncol(trash_data)` variables. The Precipitation dataset describes the amount of rain in each year, by month. It contains `r nrow(precip_data)` observations and `r ncol(precip_data)` variables. The total amount of precipiatin in 2017 was `r sum(2017$precip_data)`. The median number of sports balls in 2016 was `r 
filter(precip_data, 2016)
median(balls_integer$precip_data)`.  

#Problem 3

```{r}
library(p8105.datasets)
data(brfss_smart2010) 
brfss_smart2010 = brfss_smart2010 %>% 
  janitor::clean_names() 
```

```{r}
brfss = filter(brfss_smart2010, topic == "Overall Health") %>%  
  
  select(-class, -topic, -question, -sample_size, -confidence_limit_low, -confidence_limit_high, -display_order, -data_value_unit, -data_value_type, -data_value_footnote_symbol, -data_value_footnote, -data_source, -class_id, -topic_id, -location_id, -question_id, -respid, -geo_location) %>% 
  
  spread(response, data_value) %>% 
  janitor::clean_names() %>% 
  mutate(good_response = excellent + very_good)
```

```{r}
brfss %>% 
  distinct(locationabbr) 
#51 unique locations 
brfss %>% 
  group_by(locationabbr) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
#New Jersey is observed the most (146 times)
```

```{r}
library(lattice)

brfss %>% 
  filter(year == 2002) %>% 
  median(excellent) %>% 
  histogram(excellent, brfss)

```

```{r}
brfss %>% 
  filter by locationdesc
  ggplot(aes(x = year, y = excellent , color =  ))
```









